---
author: "Pauline"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Lab 1 - First step in Rstudio

## Cwiczenia



### CW 1

Which symbol would you use to make a comment in your code?

</ul>

<details>

<summary>ODP 1</summary>

In order to comment on your code without disturbing the rest of it, you need to use the hashtag #

</details>



### CW 2

I am creating a variable called: g <- 'Gdansk'

What is the difference between running g and print(g)?

```{r}
g <- 'Gdansk'

g

print(g)

```

<details>

<summary>ODP 2</summary>

There is no difference; both will invoke the object.

</details>



### CW 3

What are the four types of class that variables can belong to?

<details>

<summary>ODP 3</summary>

```{r}
# numeric: 
nb3.6 <- 3.6

# integer:
nb4 <- 4

# text
text <- 'Gdansk'

# factor:
color <- 'blue'

```

The four type of classes are: 
<ul>

<li>numeric: it can be an integer or a decimal number</li>

<li>integer: it can only be an integer</li>

<li>text: it can be anything, but it is usually more intuitive to combine letters to create this variable</li>

<li>factor: it can be anything, it will provide categories </li>

</ul>


</details>

### CW 4

How would you assess the class of a given variable? Can you identify the class of the following variables?

```{r}

nb3.6 <- 3.6

nb4 <- 4

text <- 'Gdansk'

color <- 'blue'

```

<details>

<summary>ODP 4</summary>

```{r}
 
is.numeric(nb3.6)

is.numeric(nb4)

is.character(text)

is.character(color)

```

The is.XXXX is returning a TRUE or FALSE response. If the response is TRUE it means that indeed the variable is from the class represented by the XXXX, otherwise it is not.

</details>

### CW 5

In the previous exercise, nb4 was classified as a numeric variable and colour as a text variable. However, we would like nb4 to be considered an integer and 'color' to be considered a factor. This can be achieved by changing the class of our object. How would you do this?

<details>

<summary>ODP 5</summary>

```{r}
# change the class of the value 4 from numeric to integer
nb4new <- as.integer(nb4)
is.integer(nb4new)

# change the class of the value 'blue' from numeric to factor
colornew <- as.factor(color)
is.factor(colornew)

```

</details>


### CW 6

In question 5, we decided to give the transformed variables different names: nb4 became nb4new, and color became colornew. What would be the consequences if we decided to keep the same names for the transformed variables?

<details>

<summary>ODP 5</summary>

In R, there is no 'undo', so if we use the same variable name, the previous variable is overwritten.

</details>

