---
author: "Kristin"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Lab 2 - Vectors and data frames

## Cwiczenia

### CW 1

We will now load a data set that is incorporated in R and look at its structure.

```{r}
str(iris)
```
This helps a lot to get an overview on what your data looks like.

Which command can you use to look at the first rows of this dataset?

<details>

<summary>ODP 1</summary>

You can use the following command. It shows you the first six rows of your data.

```{r}
head(iris)
```

</details>

### CW 2

What is the difference?

```{r}
# a
iris[2, ]

# b
iris[, 2]
```

<details>

<summary>ODP 2</summary>

The number before the comma in [x,x] refers to the row; the number behind the comma refers to the column of your data frame.

<details>

### CW 3

How to handle non existing values
When working with real data, individual values will occasionally be missing.
For example, you are measuring the wing lengths of birds, but one bird is molting or has a broken feather. You can still record the other measurements, but you are missing the wing length. 
We will simulate this here with the sepal length. Can you set the 10th value to 'NA'?

<details>

<summary>ODP 3</summary>

```{r}
iris$Sepal.Length[10] <- NA          # the 10th value is set to NA
```

<details>

### CW 4

What happens if you now calculate the mean of sepal length?

```{r}
mean(iris$Sepal.Length)
```

Can you solve this issue?

<details>

<summary>ODP 4</summary>

You can use the following command. It shows you the first six rows of your data.

```{r}
mean(iris$Sepal.Length, na.rm=TRUE)  # The NAs must be actively removed.
```

<details>

### CW 5

Let us create our own data frame (similar to the iris one).
First, create one vector for Bob, which show the times (in sec) he ran on his 100m races. He ran 8 times each.

<details>

<summary>ODP 5</summary>

```{r}
Bob <- c(62, 85, 92, 59, 68, 72, 67, 74)
```

<details>

### CW 6

Now, Eva wants to see how well she can do. Can you create another vector with her racing times?

<details>

<summary>ODP 6</summary>

```{r}
Eva <- c(59, 74, 67, 85, 67, 78, 73, 68)
```

<details>

### CW 7

Oh no, there is a mistake in the data. Correct the entry for Bob's second run.

<details>

<summary>ODP 7</summary>

```{r}
Bob[2] <- 54
```
Bob is a vector, not a data frame or matrix, so you cannot use indexing such as Bob[, 2]. R treats the vector Bob as one-dimensional, so you can only manipulate it with a single index.

<details>

