---
author: "Pauline"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Lab 6 - Using ggplot


## Cwiczenia

During this exercise session we will continue using the penguin data set and will need to download a couple of packages

```{r}
#install.packages("")
library(dplyr) ## to clean the data set => %>% , mutate(), case_when(), filter(), select()
library(tidyr) ## to tidy up the data 
library(ragg) ## to export your plots while having control over the size of the text
library(ggplot2) ## to plot
library(cowplot) ## to combine different plots together

library(palmerpenguins)
Penguins <- penguins_raw
```

Right now, the column names are not written in a efficient way for coding as they include space & parantheses. 

We will rename the column for easier data handling:

```{r}
## You can rename your colnames in different ways

## First way of doing it:
#colnames(Penguins) <- gsub(" ", ".", colnames(Penguins)) ## replace the space by a dot
#colnames(Penguins) <- gsub("\\s*\\([^\\)]+\\)", "", colnames(Penguins)) ## remove the (...)


## Second way of doing it, this is the way we will use for this training:
## Create column name that will be easier to use:
Penguins_clean <- Penguins %>% rename(Individual_ID = "Individual ID", 
                                Clutch_Completion= "Clutch Completion", 
                                Date_Egg = "Date Egg", 
                                Culmen_Length_mm = "Culmen Length (mm)",
                                Culmen_Depth_mm = "Culmen Depth (mm)",
                                Flipper_Length_mm = "Flipper Length (mm)",
                                Body_Mass_g = "Body Mass (g)",
                                Delta_15N_ooo = "Delta 15 N (o/oo)",
                                Delta_13C_ooo = "Delta 13 C (o/oo)") %>%
  # Create new variables which we will be used in our analysis:
  mutate(SEX = case_when(Sex == "MALE" ~ "M",
                         Sex == "FEMALE" ~ "F"),
         SPECIES = case_when(Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
                             Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
                             Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap"),
         Quality = case_when(Clutch_Completion == "No" ~ "Bad year",
                            Clutch_Completion == "Yes" ~ "Good year")) %>%
   # Only keep certain column:
  dplyr::select(SPECIES,SEX,Island,Quality,Date_Egg,Culmen_Length_mm,Culmen_Depth_mm,Flipper_Length_mm,Body_Mass_g,Delta_15N_ooo,Delta_13C_ooo)

```

As you can see in the last line of the code we are writing `dplyr::select` this is because we are using many different packages and there is a conflict with at least two packages using the function select. We need to specify the function from which package we are calling to avoid getting an error message.




### CW 1.1

How would you make a plot looking at the number of individual with a good year vs individual with a bad year:

</ul>

<details>

<summary>ODP 1.1 </summary>

```{r}
## Look at the number of individual with a good year vs individual with a bad year:
GoodvsBad <- Penguins_clean %>% group_by(Quality) %>% summarise(counts = n())
GoodvsBad

## Plot the results from GoodvsBad:
plot_amount1.1 <- ggplot(data = GoodvsBad, aes(x = Quality, y = counts, fill = Quality)) +
            geom_bar(stat = "identity")+
            theme_bw()
plot_amount1.1
```

</details>

### CW 1.2 

How would you make a plot looking at the number of individual with a good year vs individual with a bad year but this time you should also discriminate between species

</ul>

<details>

<summary>ODP 1.2 </summary>

```{r}
## Look at the number of individual with a good year vs individual with a bad year, for each species:
GoodvsBadSpecies <- Penguins_clean %>% group_by(Quality, SPECIES) %>% summarise(counts = n())
GoodvsBadSpecies

## Plot the results from GoodvsBadSpecies:
plot_amount1.2 <- ggplot(data = GoodvsBadSpecies, aes(x = Quality, y = counts, fill = SPECIES)) +
            geom_bar(stat = "identity", position=position_dodge())+
            theme_bw()
plot_amount1.2
```

</details>


### CW 2

Can you tell what we are representing in  plot_proportions2opt1 and plot_proportions2opt2 and tell what 
is the difference? Which part of the code is coding for the difference ?

```{r}
head(Penguins_clean)

GoodvsBadSpecies_perc <- Penguins_clean %>% group_by(Quality, SPECIES) %>% 
  summarise(counts = n()) %>% 
  mutate(perc_GoodBadSex=round(prop.table(counts) * 100))
GoodvsBadSpecies_perc

plot_proportions2opt1 <- ggplot(data = GoodvsBadSpecies_perc, aes(x = Quality, y = perc_GoodBadSex, fill = SPECIES)) +
            geom_bar(stat = "identity", position=position_dodge())
plot_proportions2opt1

plot_proportions2opt2 <- ggplot(data = GoodvsBadSpecies_perc, aes(x = Quality, y = perc_GoodBadSex, fill = SPECIES)) +
            geom_bar(stat = "identity")
plot_proportions2opt2

## This piece of code help you to have both plot next to each other:
library(cowplot)
plot_grid(plot_proportions2opt1,plot_proportions2opt2)

```

<details>

<summary>ODP 2</summary>

Both plot are looking at if the year was good or bad for each penguins species and is expressing this result in percentage. 
The difference between the two code is found in the command `position=position_dodge()`. In plot_proportions1, the bars are standing next to one another whereas in plot_proportions2, the bars are stacked above each other


</details>



### CW 3.1

How would you plot the graph to look at the body mass for Adelie Penguin only?

<details>

<summary>ODP 3</summary>

```{r}

## Solution 1:
plot_distribution3.1opt1 <- ggplot(subset(Penguins_clean, SPECIES == "Adelie"), aes(Body_Mass_g,fill = SPECIES, colour = SPECIES)) +
  geom_density(alpha = 0.5)
plot_distribution3.1

## Solution 2:
Adelie <- Penguins_clean %>% filter(SPECIES == "Adelie")

plot_distribution3.1opt2 <- ggplot(Adelie, aes(Body_Mass_g, fill = SPECIES, colour = SPECIES)) +
  geom_density(alpha = 0.5)
plot_distribution3.1opt2
```
</details>


### CW 3.2

How would you plot the graph to look at the body mass for Adelie Penguin only but discriminate between sex?

<details>

<summary>ODP 3.2</summary>

```{r}
## If you forget to remove the NA, you will have the following plot:
Adelie <- Penguins_clean %>% filter(SPECIES == "Adelie")

plot_distribution3.2opt1 <- ggplot(Adelie, aes(Body_Mass_g,fill = SEX, colour = SEX)) +
  geom_density(alpha = 0.5)
plot_distribution3.2opt1


## You first need to remove the NA, otherwise they will appears in your plot
AdelieSex <- Penguins_clean %>% filter(SPECIES == "Adelie" & !is.na(SEX))

## This plot shows the body mass for Adelie Penguin with one curve for females and one curve for males:
plot_distribution3.2opt2 <- ggplot(AdelieSex, aes(Body_Mass_g,fill = SEX, colour = SEX)) +
  geom_density(alpha = 0.5)
plot_distribution3.2opt2

## This shows you both plot next to each other:
library(cowplot)
plot_grid(plot_distribution3.2opt1,plot_distribution3.2opt2)


```
</details>

### CW 4.1

How would you look at the realtionship between Delta_13C and Detla_15N?

<details>

<summary>ODP 4.1 </summary>

```{r}
## First we need to remove the NA for the variable Delta15N & for Delta13C:
Penguins_delta <- Penguins_clean %>% filter(!is.na(Delta_13C_ooo)) %>% filter(!is.na(Delta_15N_ooo))

## We can look at the relation ship between Delta15N & Delta13C
plot_xyrelationship1 <- ggplot(Penguins_delta, aes(x = Delta_15N_ooo, y = Delta_13C_ooo)) +
  geom_point(size = 1, shape = 21)
plot_xyrelationship1
```

</details>


### CW 4.2

Actually, you can have fun changing the shape, size and color of the points from the previous plot, using the following link:

<ul>   
<li> color: https://www.sthda.com/english/wiki/colors-in-r </li>
<li> shape: https://stackoverflow.com/questions/46803260/assigning-40-shapes-or-more-in-scale-shape-manual </li>
<li> size: https://www.statology.org/ggplot-point-size/#:~:text=by%20Zach%20Bobbitt%20Last%20updated,y=wt))%20+%20geom_point() </li>
</ul>

Try and change the size, color and shape of the points based on the Species

```{r}
head(Penguins_delta)

## We can look at the relation ship between Delta15N & Delta13C
plot_xyrelationship4.2 <- ggplot(Penguins_delta) +
  geom_point(aes(x = Delta_15N_ooo, y = Delta_13C_ooo, size = SPECIES, color = SPECIES, shape = SPECIES)) +
  scale_size_manual(values=c(2,3,4)) +
  scale_color_manual(values = c("pink4","purple3","darkblue")) +
  scale_shape_manual(values=c(11,8,25))
plot_xyrelationship4.2
```


<details>

<summary>ODP 4.2 </summary>

```{r}

### Here is an example of how the parameters where changed:
plot_xyrelationship4.2 <- ggplot(Penguins_delta) +
  geom_point(aes(x = Delta_15N_ooo, y = Delta_13C_ooo, size = SPECIES, color = SPECIES, shape = SPECIES)) + # Here you are stating that the size, the color and the shape of the points should be dependent of the species of penguin
  scale_size_manual(values=c(4,4,4)) + # because we have three species in our data set, we need to provide three values 
  scale_color_manual(values = c("dodgerblue3","lightblue4","cadetblue")) + # because we have three species in our data set, we need to provide three values 
  scale_shape_manual(values=c(97,99,103)) # because we have three species in our data set, we need to provide three values 
plot_xyrelationship4.2
```

</details>

### CW 5

How would you calculate standard deviation around the mean for the flipper length of all penguin species?

<details>

<summary>ODP 5 </summary>

```{r}
head(Penguins_clean)

## We will calculate the standard deviation around the mean for the flipper length of all penguin species:
Flipper_Species <- Penguins_clean %>% group_by(SPECIES) %>% 
  summarise(mean_flipper = mean(Flipper_Length_mm, na.rm=T),
            sd_flipper = sd(Flipper_Length_mm, na.rm=T))


## Here is a way to vizualize those results using ggplot:
## Visualize the difference in flippers length for each penguin species:
plot_uncertainty5 <- ggplot(Flipper_Species, aes(x = SPECIES, y = mean_flipper, color = SPECIES)) +
  geom_pointrange(aes(ymin=mean_flipper-sd_flipper, ymax=mean_flipper+sd_flipper))
  
plot_uncertainty5
```
</details>


### CW 6 - Not for the faint hearted ;) 

This part is more to show you how we can build complex plot with ggplot by stacking different layers together. 
You can copy this code in you Rstudio and try and play with the parameters of the `theme()` section to see how it affect the final plot


```{r}

## Remove all the NA from the data set:
Penguins_allnarm <- Penguins_clean %>%
  na.omit()

## Reformat my table:
library(tidyr) ## required to use the function gather()
Penguinslong <- gather(Penguins_allnarm, key="measure", value="value", c("Culmen_Length_mm", "Culmen_Depth_mm","Flipper_Length_mm","Body_Mass_g"))
Penguinslong$measure <- as.factor(Penguinslong$measure)

## Rename my variables for them to have a pretty name appearing in the final plot:
Penguinslong <- Penguinslong %>% mutate(measure = case_when(measure == "Culmen_Length_mm" ~ "Culmen length (mm)",
                                                            measure == "Culmen_Depth_mm" ~ "Culmen depth (mm)",
                                                            measure == "Flipper_Length_mm" ~ "Flipper length (mm)",
                                                            measure == "Body_Mass_g" ~ "Body mass (g)"))




## Build the plot:
Penguins_SpeciesIsland_plot <- ggplot(Penguinslong, aes(x = as.factor(SPECIES), y = value)) +
  # Individual points
  geom_point(aes(color = SPECIES),position = position_jitter(w = .15), alpha=0.5) +
  scale_color_manual(values = c("dodgerblue3","lightblue4","cadetblue")) +
  # Violon plot: distributions 
  geom_violin(aes(fill = SPECIES), alpha = 0.5) + # geom_flat_violin would not accept scale free from facet_grid or facet_wrap
  scale_fill_manual(values = c("dodgerblue3","lightblue4","cadetblue")) +
  # Boxplot: distributions
  geom_boxplot(aes(fill = SPECIES),width = .25,outlier.shape = NA) +
  # Create the facet plot: combine different variables all together on the same plot
  facet_grid(measure ~ Island, scales = "free", #  Faceting with free y-scales to allow individual y-ranges for each row
             switch = "y") +  # switch the side of the measured variables names from right to left
  theme_bw() +  # Optional, better visualization with a clean background
  # Adjust every parameter in your plot as you like
  theme(
    strip.text.x = element_text(size = 12),  # Adjust text size for Species facets
    strip.text.y = element_text(size = 12),   # Adjust text size for measure facets
    axis.text = element_text(size=12, color = "black"), # Adjust the axis text
    axis.title = element_blank(),  # Adjust the axis title
    legend.title = element_text(size = 12), # Adjust the legend title 
    legend.text = element_text(size = 12) # Adjust the legend text
  ) +
  scale_y_continuous("value", position = "right")  # change the y-axis scale on the right
Penguins_SpeciesIsland_plot


### Try to export the plot directly from the console while having control over the size of the text and lines:
setwd("")
library(ragg)
## Scaling change the size of the text on my plot
ragg::agg_png("Penguins_SpeciesIsland_plot_10x10_2.png", width = 10, height = 10, units = "in", res = 300, scaling = 2)
Penguins_SpeciesIsland_plot
dev.off()


```





