---
author: "Kristin"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Lab xxx

## Cwiczenia - histograms in ggplot

### CW 1

We will now import a data set that contains the data for the following exercises.
Can you import the data called 'birth_places_uni.csv'?

<details>

<summary>ODP 1</summary>

You can use the following command but will need to replace the working directory with your folder names of where you are saving your files:

```{r}
birth_places_uni <- read.csv("C:/Users/krist/Desktop/birth_places_uni.csv")
```

Another way to do it is using the button 'session', selecting the working directory, going to the tab files, and uploading it from there.

</details>

### CW 2

Can you look at the structure of the data?
What is wrong? Does something need to be changed/adjusted?

```{r}
str(birth_places_uni)
```

<details>

<summary>ODP 2</summary>

We will need to change 'Sex' and 'Distance' to a factor and numerical, respectively.

<details>

### CW 3

Let's change 'Sex' and 'Distance' to what we need to continue.

<details>

<summary>ODP 3</summary>

```{r}
birth_places_uni$Sex <- as.factor(birth_places_uni$Sex)

birth_places_uni$Distance.to.University..km. <- as.numeric(birth_places_uni$Distance.to.University..km.)

```

<details>

### CW 4

Can you create a histogram using ggplot that we got to know during class?

<details>

<summary>ODP 4</summary>

```{r}
library(ggplot2)
ggplot(birth_places_uni, aes(x = Distance.to.University..km.)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +  # Set bin width
  theme_minimal()

```
As you can see, we can change the colors.

<details>

### CW 5

Now, we do not have very nice axis descriptions. Can you change them to make it a better plot?

<details>

<summary>ODP 5</summary>

```{r}
ggplot(birth_places_uni, aes(x = Distance.to.University..km.)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
labs(title = "Histogram of Distances to University", 
       x = "Distance to University (km)", 
       y = "Frequency") +
  theme_minimal()
```

<details>

### CW 6

Do you know what this plot shows? What is different in the code from the previous code?

```{r}
ggplot(birth_places_uni, aes(x = Distance.to.University..km.)) +
  geom_histogram(binwidth = 5, fill = "skyblue", color = "black") +
labs(title = "Histogram of Distances to University", 
       x = "Distance to University (km)", 
       y = "Frequency") +
  facet_wrap(~Sex) +
  theme_minimal()
```

<details>

<summary>ODP 6</summary>

We firstly added a part of the code that creates two separate plots for male and female students and their respective distances between their place of birth and university. 

<details>

### CW 7

Play with the 'binwidth'. What changes in your plot?

<details>

<summary>ODP 7</summary>

Here, we changed the binwidth for better visualization. Now we can better see the differences between the distances of males and females.

```{r}
ggplot(birth_places_uni, aes(x = Distance.to.University..km.)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
labs(title = "Histogram of Distances to University", 
       x = "Distance to University (km)", 
       y = "Frequency") +
  facet_wrap(~Sex) +
  theme_minimal()
```

<details>

