---
author: "Pauline"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Vizualisation of Geospatial data

```{r}
#install.packages("")
library(dplyr) ## to clean the data set => %>% , mutate(), case_when(), filter(), select()
library(tidyr) ## to tidy up the data 
library(lubridate) ## to play with date and time format
library(ragg) ## to export your plots while having control over the size of the text
library(ggplot2) ## to plot
library(esquisse) ## if you're new to R and find creating plots with code intimidating, this package is for you!
library(ggridges) ## extension needed in ggplot when stacking density plots
library(cowplot) ## to combine different plots together
library(sf) ## to use certain function when playing with geospatial data
library(raster) ## to download raster files in R
library(rnaturalearth) ## to have the land polygon
library(ggspatial) ## to plot geospatial data with ggplot
library(viridis) ## to use specific color gradient in ggplot
library(stars) ## to tidy up geospatial data

library(palmerpenguins)
Penguins <- penguins_raw
```

Right now, the column names are not written in a efficient way for coding as they include space & parantheses. 

We will rename the column for easier data handling:

```{r}
## You can rename your colnames in different ways

## First way of doing it:
#colnames(Penguins) <- gsub(" ", ".", colnames(Penguins)) ## replace the space by a dot
#colnames(Penguins) <- gsub("\\s*\\([^\\)]+\\)", "", colnames(Penguins)) ## remove the (...)


## Second way of doing it, this is the way we will use for this training:
## Create column name that will be easier to use:
Penguins_clean <- Penguins %>% rename(Individual_ID = "Individual ID", 
                                Clutch_Completion= "Clutch Completion", 
                                Date_Egg = "Date Egg", 
                                Culmen_Length_mm = "Culmen Length (mm)",
                                Culmen_Depth_mm = "Culmen Depth (mm)",
                                Flipper_Length_mm = "Flipper Length (mm)",
                                Body_Mass_g = "Body Mass (g)",
                                Delta_15N_ooo = "Delta 15 N (o/oo)",
                                Delta_13C_ooo = "Delta 13 C (o/oo)") %>%
  # Create new variables which we will be used in our analysis:
  mutate(SEX = case_when(Sex == "MALE" ~ "M",
                         Sex == "FEMALE" ~ "F"),
         SPECIES = case_when(Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
                             Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
                             Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap"),
         Quality = case_when(Clutch_Completion == "No" ~ "Bad year",
                            Clutch_Completion == "Yes" ~ "Good year")) %>%
   # Only keep certain column:
  dplyr::select(SPECIES,SEX,Island,Quality,Date_Egg,Culmen_Length_mm,Culmen_Depth_mm,Flipper_Length_mm,Body_Mass_g,Delta_15N_ooo,Delta_13C_ooo)

```

As you can see in the last line of the code we are writing `dplyr::select` this is because we are using many different packages and there is a conflict with at least two packages using the function select. We need to specify the function from which package we are calling to avoid getting an error message.

We can combine spatial data with remote sensing, in our case, we are combining it with the data coming from Copernicus from the following link: https://data.marine.copernicus.eu/product/SST_GLO_SST_L4_NRT_OBSERVATIONS_010_001/description

To understand the data you are extracting and are about to analyze, it is helpful to check the user manual provided on the product's webpage.

In the User Manual we can see that the analysed_sst in the file is express in Kelvin 

We can visualize spatial data using 
```{r}
head(Penguins_clean)

### Coordinates of the three colonies: Latitude then Longitude
Dream <-c(-64.730419, -64.247602)
Biscoe <- c(-66.286626, -67.033080) 
Torgersen <- c(-64.765938, -64.082502)

Colonies <- data.frame(Names =c("Dream","Biscoe","Torgersen"),
                       Latitude = c (-64.730419, -66.286626,-64.765938),
                       Longitude = c(-64.247602,-67.033080,-64.082502))

## Extract the land map of the area we are interested in
library(rnaturalearth)
antarctica <- ne_countries(scale = 10, returnclass = "sf", country = "Antarctica")
plot(antarctica)

## Extract only the area I want to focus on which is the Region of Anvers where we have the three island:
library(sf)
anvers <- st_crop(antarctica$geometry, xmin = -70.4, xmax = -63, ymin = -66.5, ymax = -64.5) 
plot(anvers)

### Import data remote sensing data:
library(raster)
SST <- raster("~/YOUR_COMPUTER_DIRECTORIES/20251214120000-UKMO-L4_GHRSST-SSTfnd-OSTIA-GLOB-v02.0-fv02.0.nc")


### Create a crop to extract the data only for the area of interest
pr.e <- extent(-70.4, -63, -66.5, -64.5) # Anvers Region
pr.SST <- crop(SST, pr.e)
dfSST <- as.data.frame(pr.SST, xy = TRUE)
str(dfSST)

### Convert the data from Kelvin to °C:
dfSSTc <- dfSST %>% mutate(SST_celsius = analysed.sea.surface.temperature - 273.15) %>%
  mutate(Zone = case_when(SST_celsius < 0 ~ "Very cold",
                          SST_celsius >= 0 ~ "Cold"))
  
range(dfSSTc$SST_celsius, na.rm=T)
unique(dfSSTc$SST_celsius, na.rm=T)


### We can plot everything the first results: including the Zone category color, the colonies locations, and the map of the island
library(ggspatial) ## necessary to use geom_spatial_point()
plot_geospatial1 <- ggplot() +
  geom_raster(data = dfSSTc, aes(x = x, y = y, fill = as.factor(Zone))) +
  scale_fill_manual(name = "Sea Surface Zones", values = c("#33CCFF","#FFFFFF")) +
  # Show the map:
  geom_sf(data = anvers, fill = "bisque3", inherit.aes = FALSE) +
  # Show the location of the three different colonies:
  geom_spatial_point(data=Colonies,aes(x=Longitude,y=Latitude),
                     size=3, crs=4326, shape = 21, stroke = 0.5, fill = c("#FF33CC")) +
  labs(title = "Colonies and SST")
plot_geospatial1

### We can plot everything the first results: including the Zone category color, the colonies locations, and the map of the island
library(viridis) ## necessry to use scale_fill_viridis
plot_geospatial2 <- ggplot() +
  geom_raster(data = dfSSTc, aes(x = x, y = y, fill = SST_celsius)) +
  scale_fill_viridis(option = "D") + ## you can try different letter from A to H
  # Show the map:
  geom_sf(data = anvers, fill = "bisque3", inherit.aes = FALSE) +
  # Show the location of the three different colonies:
  geom_spatial_point(data=Colonies,aes(x=Longitude,y=Latitude),
                     size=3, crs=4326, shape = 21, stroke = 0.5, fill = c("#FF33CC")) +
  labs(title = "Colonies and SST")
plot_geospatial2

```

If you work with GPS tracking data, you can extract the information from remote sensing for the specific GPS location and you can plot everything together


```{r}
## Let's create fake GPS position in the Anvers Region:
bbox_lonlat <- st_bbox(c(xmin = -70.4, xmax = -63, ymin = -66.5, ymax = -64.5),
                       crs = st_crs(4326))

## Convert the box into a polygon:
bbox_poly <- st_as_sfc(bbox_lonlat)

## Sample in the area of interest
set.seed(123) ## By setting the seed we are always creating the same random point
sample_points <- st_sample(bbox_poly, size = 200)

## Remove sampled point thats might have been created on land that are in the area:
random_GPS_points <- sample_points[
  !st_intersects(sample_points, anvers, sparse = FALSE)
]

## keep only 30 computed random points
random_GPS_points30 <- random_GPS_points[1:30]
head(random_GPS_points30)
random_GPS_points30df <- as.data.frame(st_coordinates(random_GPS_points30))

## We randomly assign the GPS point to a specific penguin species, by taking the X value as the classifying parameter:
random_GPS_points30df <- random_GPS_points30df %>% mutate(Species = case_when(X < -68 ~ "Adelie",
                                                                              X > -69 & X < -64 ~ "Gentoo",
                                                                              X > -65 ~ "Chinstrap"))

## Visually inspect if the created GPS position have correct coordinates and overlap with the study area:
plot_geospatial3 <- ggplot() +
  geom_raster(data = dfSSTc, aes(x = x, y = y, fill = as.factor(Zone))) +
  scale_fill_manual(name = "Sea Surface Zones", values = c("#33CCFF","#FFFFFF")) +
  # Show the map:
  geom_sf(data = anvers, fill = "bisque3", inherit.aes = FALSE) +
  # Show the location of the three different colonies:
  geom_spatial_point(data=Colonies,aes(x=Longitude,y=Latitude),
                     size=3, crs=4326, shape = 21, stroke = 0.5, fill = c("#FF33CC")) +
  ## Show the location of the fake-GPS points:
  #geom_point(data = random_GPS_points30df, aes(x = X, y = Y), color = "black", shape = 21, size = 2, fill = c("#FFCCCC")) +
  geom_point(data = random_GPS_points30df, aes(x = X, y = Y,  color = Species), size = 2) + ## if you want to color by Species of penguins
  labs(title = "Colonies and SST")
plot_geospatial3


### Add the SST to the specific GPS positions:
library(stars)
## Convert the SST raster object to stars format:
starsSST <- st_as_stars(pr.SST, crs = st_crs(4326))
## Make sure the GPS object is in the same format as the SST object
random_GPS_points30dfsf <- st_as_sf(random_GPS_points30df, coords = c("X", "Y"), crs = 4326)

## Extract cell values (SST values) at point locations (GPS locations)
SSTextractGPS <- st_extract(starsSST, random_GPS_points30dfsf) %>% st_as_sf() 

## Combine the GPS position where we have the penguins information with the extracted SST info for those GPS positions
GPSxySST <- cbind(random_GPS_points30dfsf, SSTextractGPS) 

## Convert the column geometry in coordinates X and Y then extract it
GPSxySSTcoord <- st_coordinates(GPSxySST)
## Attach the coordinates in the X and Y format to our table
GPSSSTcombined <- cbind(GPSxySST, GPSxySSTcoord)
GPSSSTcombined
```
